# Results {#results}
We ran 10 different ActivitySim-to-BEAM mode choice combinations and processed their ride hail statistics and service capabilities into results. Specifically, we obtained ridership (Section \@ref(res-ridership)) and wait time (Section \@ref(res-waits)) information from the final iteration of BEAM for each scenario. Alongside each other, these results should shed some light on how differing mode choice structures between an activity-based model and a multi-agent simulation affect the forecasting of novel modes.


## Ridership {#res-ridership}
As novel modes, like ride hail, become more popular it is ever more important to accurately predict ridership. Predicting accurate ridership totals will allow for efficient driver allocation, decreased traffic congestion, and improved rider experience [@chen21]. Table \@ref(tab:ridership) shows the ridership percentages of ride hail, pooled ride hail, and ride hail transit modes for all 10 mode choice combinations. Correctly understanding the results of Table \@ref(tab:ridership) may help forecasters predict ridership numbers more effectively in the future. To accurately interpret the results of Table \@ref(tab:ridership) however, we must determine the best method to reading the table. The many slight variations in each mode choice structure makes interpreting the results complicated.

```{r ridership, echo=FALSE}
ridership <- tar_read(ridership)
kable(ridership, caption = 'Percent ride hail ridership by mode choice combination scenario.', booktabs = TRUE)  %>%
  kable_styling(latex_options="scale_down")
```

We propose two methodologies to interpreting the results of Table \@ref(tab:ridership) as well as the other results presented in Section \@ref(res-waits). 

  1. Comparing scenarios individually
  2. Comparing scenarios collectively
  
###  Comparing Scenarios Individually {#res-ind}
To provide an example of using the first methodology, we compare the AsimRideHail scenario with the other nine scenarios. To begin, the AsimRideHail scenario compared with the NoRideHail scenario helps us understand how ActivitySim estimates ridership. Since the AsimRideHail scenario and NoRideHail scenario are run with BEAM's modal innovation turned off, we can assume that these results mimic those output from ActivitySim. Obviously the NoRideHail scenario will produce no ride hail results. Looking at the AsimRideHail scenario though, we clearly see the level at which ActivitySim predicts ridership. ActivitySim estimates that of the entire population's mode choice, 0.0024% are ride hail modes, 0.003% are pooled ride hail modes, and 0.000% are ride hail transit modes. Since 0.000% are ride hail transit trips, we conclude that ActivitySim may struggle at creating this particular mode option. 

Comparing the AsimRideHail scenario with the *RideHail* type scenarios (BeamRideHail and AsimBeamRideHail), less direct conclusions can be made. Of course, we obviously see that the ridership totals for all ride hail modes in all four scenarios are greater than the ridership of the AsimRideHail scenario. We can conclude that the BEAM mode choice structure which allows only ride hail modal innovation will predict higher ridership totals than ActivitySim. However, since each scenario presents a unique total ridership for each ride hail mode, we cannot draw any more conclusions by only referencing the results of the AsimRideHail scenario. 

Comparing the AsimRideHail scenario with the BeamAll scenarios we see the effectiveness of BEAM at estimating ride hail modes without ride hail being selected in the initial plans. Both BeamAll scenarios display a large increase in total ride hail ridership when compared with the AsimRideHail scenario. Comparing the *All* type scenarios (BeamAll and AsimBeamAll), we notice that BEAM will estimate higher ridership totals than ActivitySim, independent of whether or not the input plans include ride hailing modes. Comparing the AsimRideHail scenario with the BeamAll:Path and AsimBeamAll:Path scenarios we see a glimpse at how BEAM estimates ride hail modes with only path variables, independent of whether or not the input plans include ride hail. A significant increase in ridership in these two scenarios is displayed when compared with the AsimRideHail scenario. This may highlight the importance of including path variables to estimate ride hail ridership. 

Finally, comparing the AsimRideHail scenario with the AsimBeamAll:PPL scenario we learn about the inter-workings of the BEAM mode choice that is most consistent with ActivitySim. The BeamAll:PPL and AsimBeamAll:PPL scenarios reference a BEAM mode choice structure which uses all modal options and all utility parameters. In other words, this BEAM mode choice structure is the one most aligned with that of ActivitySim. By comparing AsimBeamAll:PPL with the AsimRideHail scenario, we see how ride hail statistics *differ* when modeled with an activity-based model than with a multiagent simulation. Looking at the differences between the the AsimRideHail and AsimBeamAll:PPL scenarios in Table \@ref(tab:ridership), we see an increase of 0.256% of total ride hail modes, 0.533% of total pooled ride hail, and 0.014% of total ride hail transit modes. Overall, a significant increase in ridership was achieved by simply turning on modal innovation in BEAM. Overall, we conclude that BEAM is more prone to estimating higher ridership totals for ride hailing modes than ActivitySim.

Comparing other scenarios between each other individually may help us draw more conclusions, especially with the *RideHail* type scenarios, however, analyzing every combination of individual scenario is beyond the extent of this paper. As a result, we instead move to comparing these scenarios collectively to better understand the ridership results. 

### Comparing Scenarios Collectively {#res-col}
Looking at scenarios collectively, and comparing them with each other and with other collective groups will extract deeper results in our research. Section \@ref(meth-scenarios) and Table \@ref(tab:tbexperiments) describe the following three different descriptors which will be used as the collective groups:

  1. ActivitySim Mode Choice / Input Plans
  2. BEAM Mode Choice
  3. BEAM Utility Variables

As shown in Table \@ref(tab:ridership), minimal differences in ridership is produced between ActivitySim mode option outputs (scenarios with the *Asim* prefix vs. without). For example, comparing the *RideHail* type scenarios we see no direct correlation with ridership and input plans. With  the BeamRideHail:Path and AsimBeamRideHail:Path, we see that when ride hail doesn't exist in the input plans, the total ride hail type modes (ride hail, pooled ride hail, and ride hail transit) complete about 6.2% of total modal distribution whereas when ride hail type modes are included, total ride hail type modes complete about 5.8% of total modal distribution. Although both project an enormous share using ride hail, the scenario with ride hail in the input plans projects more ride hail ridership. Oddly enough, however, when comparing the BeamRideHail:PPL and AsimBeamRideHail:PPL scenarios, the scenario with ride hail in the input plans projects less ride hail ridership. In these two scenarios we see that when ride hail doesn't exist in the input plans, the total ride hail type modes complete about 5.3% of total modal distribution whereas when ride hail type modes are included, total ride hail type modes complete around 6.6% of total modal distribution. Overall, no direct correlation between ride hail ridership totals and input plans is seen between *RideHail* type scenarios. Then, when examining BeamAll:PPL and AsimBeamAll:PPL, the output ridership percentages are almost identical. When using the BEAM mode choice model that is most consistent with that of ActivitySim, BEAM estimates ride hail ridership at the same level with or without ride hail in the input plans. However, when examining BeamAll:Path and AsimBeamAll:Path we see that when only path variables are used in the utility equation, ridership is not identical between ActivitySim mode options. BeamAll:Path shows an increase in ridership for ride hail modes and a decrease in pooled ride hail and ride hail transit modes than in AsimBeamAll:Path. Initial plan sets play a role in who selects pooled and transit ride hail trips when only using path variables to determine mode choice. We hypothesize that with the lack of person attributes in the utility equation, pooled and transit ride hail options look less appealing. Collectively, comparing Scenarios AsimRideHail, AsimBeamRideHail:Path, AsimBeamRideHail:PPL, AsimBeamAll:Path, and AsimBeamAll:PPL against Scenarios NoRideHail, BeamRideHail:Path, BeamRideHail:PPL, BeamAll:Path, and  BeamAll:PPL we see similar results among similar mode choice structures. Although minimal shifts in total ridership were noticed in most cases, as a whole, we determine that changing the ActivitySim mode options has no *significant* change in final ride hail ridership; at least not a traceable change. 

However, when analyzing the BEAM mode options from Table \@ref(tab:ridership) as described in Table \@ref(tab:tbexperiments), obvious results are seen. The *None* type scenarios (NoRideHail and AsimRideHail) produce little to no agents choosing ride hail. The reasoning behind this is described in Section \@ref(res-ind). On the other hand, the *RideHail* type scenarios (BeamRideHail and AsimBeamRideHail) produce the largest number of ride hail modes amoung BEAM mode choice structures! This is because with the *RideHail* type mode choice, walking individuals only have the option to continue walking or take ride hail. Taking a ride hail mode is usually more attractive than walking a long distance; and because of this limitation, an excessive number of agents switch to using ride hail. Finally, the *All* type scenarios (BeamAll and AsimBeamAll) produce a significantly larger number of ride hail modes than the *None* type. According to BEAM, ride hail type modes are more attractive than initially projected by ActivitySim. Ridership is affected significantly by which mode choice structure is used by BEAM; this conclusion is clear. 

Finally, we can analyze the effect the BEAM utility variables have on ride hail ridership totals. Two different options are shown: Path and PPL. To best understand the effect the utility variables have, we compare similar BEAM structures with the utility variables adjusted. For example, BeamRideHail:Path and BeamRideHail:PPL can be compared together. In this case, the ride hail ridership decreases (3.814% to 1.615%) when using path, person, and location variables, but increases with pooled ride hail (2.120% to 3.327%) and ride hail transit (0.249% and 0.395%) when using path, person, and location variables. This same pattern occurs when analyzing the difference between the BeamAll:Path and BeamAll:PPL, and AsimBeamAll:Path and AsimBeamAll:PPL scenarios. For some oddity, AsimBeamRideHail:Path and AsimBeamRideHail:PPL follow an opposite pattern where ride hail increases (1.824% to 3.980%) and ride hail pooled decreases (3.465% to 2.293%) and ride hail transit decreases (0.461% to 0.323%) when using path, person, and location variables. Similarly, with all comparisons except between BeamRideHail:Path and BeamRideHail:PPL, we see that total ride hail percentages increases when using all path, person, and location variables to estimate ride hail. We acknowledge that not all scenarios follow the same pattern, but conclude that in general, using only path variables to estimate ride hail ridership will result in less total ride hail ridership than if using all path, person, and location type variables.

### Summary

Analyzing ridership among different mode choice structures has highlighted some important conclusions:

  - Ride hail is affected significantly by which mode choice structure is used by BEAM.
    
    - The *RideHail* BEAM mode structure will estimate much more ride hail usage than the other versions.
    
    - The *All* BEAM mode structure is prone to estimating higher ride hail usage  than ActivitySim.
 
  - Initial plans, and whether or not they include ride hail, do not significantly affect the level at which BEAM estimates ride hail. 
  
  - Path variables affects ridership; and using all path, person, and location type variables will increase total ridership.
  
  - With the lack of person attributes in the utility equation, pooled and transit ride hail options look less appealing.

By analyzing wait times with a similar approach, individually and collectively, many of these conclusions are verified.

## Wait Times {#res-waits}
Wait time is one of the most important factors people use when determining their travel mode. When traveling, long wait times are seen as burdensome; short wait times are also seen as burdensome when compared to modes with zero wait time. Therefore, estimating accurate wait times would be helpful to forecasters in understanding the usage of novel modes. Ride hail, in particular, is an up and coming novel mode where wait time definitely plays a factor in its usage. According to @dong20, wait time for ride hailing modes seems less burdensome than wait time for transit modes. This behavior may lead to higher ridership for ride hail vehicles than for transit vehicles. Since ridership hinges on wait time among other things, the accurate estimation of wait times (and other service capabilities) in essential to forecasting novel modes like ride hail.

Figure \@ref(fig:waits) shows the results of wait times for ride hail vehicles for the 10 scenarios that were run. (The NoRidehail scenario, is not shown because no ride hail vehicles were present in that scenario). By analyzing the wait times with violin plots, we analyze a detailed distribution of the ride hail wait times for each scenario.

```{r waits, echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "Distribution of ride hail wait times by mode choice combination scenario.", fig.align='center'}
tar_read(waits)
```

### Comparing Scenarios Individually
As was done in Section \@ref(res-ind), we compare some scenarios individually with each other. To start, we look at the differences that exist between AsimRideHail and AsimBeamAll:PPL. As explained when looking at ridership, this comparison explains the inter-workings of the BEAM mode choice model that is most consistent with ActivitySim. As ridership experiences an increase between these two scenarios, wait times also increase when using the BEAM mode choice model (minimum wait time increases from 0.7 to 1 minute, and mean wait time increases from 7.8 to 9.3 minutes). A noticeable increase in wait time was achieved by simply turning on modal innovation in BEAM. Since higher wait time suggest higher ridership, we again conclude that BEAM is more prone to estimating more ride hail usage than ActivitySim.

Comparing AsimRideHail and BeamAll:PPL we see the lack of impact that  ride hail in the initial plans has on effective ride hail usage. The wait time mean may have increased from 7.8 to 8.4 minutes, but the maximum wait time decreased from 20.8 to 18.3 minutes! Although no ride hail existed in the input plans, BEAM was effectively able to calculate the usage of more ride hail vehicles with less maximum wait times. Comparing AsimRideHail with BeamAll:Path and AsimBeamAll:Path, we can learn of the effect using only Path variables has on wait time. In both BeamAll:Path and AsimBeamAll:Path, mean wait time increases when compared with AsimRideHail, but AsimBeamAll:Path displays a larger maximum wait time than BeamAll:Path and AsimRideHail. This may highlight the importance of also using person and location variables to estimate ride hail wait time. The individual comparisons between AsimRideHail and the *RideHail* type scenarios are not included in this analysis, but can be personally examined using the same technique. 

### Comparing Scenarios Collectively
As with ridership, analyzing wait times on a collective scale sheds additional light in the effect mode choice has on ride hail service capabilities. To start we analyze the effect the input plans have on ride hail wait times. Collectively, comparing the scenarios with the *Asim* prefix against the scenarios without, we see a slight difference in maximum wait times. For example, Scenarios AsimBeamRideHail:PPL, AsimBeamAll:Path, and AsimBeamAll:PPL have higher maximum wait times when compared to their counterparts in BeamRideHail:PPL, BeamAll:Path, BeamAll:PPL. Although this is true, when comparing the mean values of those same scenarios, relatively no difference exists. Its hard to conclude the exact effect that having ride hail in the input plans has, but we propose a hypothesis. We suppose that the existence of ride hail in the initial plans will not affect *most* ride hail wait times. With the exception of the maximum value, similar wait time distributions and means exist between different input ride hail wait times. The presence of ride hail in the input plans may, however, affect those few riders who start with ride hail in their input plans; maybe they are the users with longer wait times. There is insufficient data to conclude this though, and even so, the majority of wait time calculations are not affected. As concluded by analyzing ridership values, as a whole we determine that changing the ActivitySim mode options creates insignificant changes in the final ride hail performance in BEAM.

Alternately, comparing different BEAM mode options does significantly affect ride hail wait times. The *None* type scenario (AsimRideHail) has the largest spread of wait times, the lowest mean wait time, and is "bottom heavy" -- referring to the fact that a major cluster of users wait less than 7.5 minutes. The *All* type scenarios have higher mean wait times (~9.3 minutes) than the *None* type and lower mean wait times than the *RideHail* type. Neither top nor bottom heavy, the *All* type scenarios seem to have a more even spread in wait times, ranging from 0.4 to 21 minutes (depending on the exact mode choice structure). BEAM seems to paint ride hail alternatives as more desirable than ActivitySim, as more users are willing to wait longer (12 to 18 minutes in the *All* scenarios). This is especially true with the *RideHail* type scenarios, as these scenarios are the most "top heavy" in comparison; a large cluster of users are willing to wait 7.5 to 20 minutes. The *RideHail* type scenarios have the largest mean wait times (~10.65 minutes). Overall, wait time is significantly affected by which mode choice structure is used by BEAM, just like as was concluded with ridership.

The last group to compare collectively is between the Path and PPL models. By comparing BeamRideHail:Path and BeamRideHail:PPL, BeamAll:Path and BeamAll:PPL, and AsimBeamAll:Path and AsimBeamAll:PPL, we see that the Path models estimate a slightly higher maximum wait time. In addition, BeamRideHail:Path and AsimBeamRideHail:Path seem to have a larger cluster above a 10 minute wait time than BeamRideHail:PPL and AsimBeamRideHail:PPL. Besides these two observations though, the differences between utility parameters is minimal Although ridership was affected by which utility parameters were used, wait time is only slightly affected. We conclude that by lacking person and location variables, there is potential for a longer wait time for ride hail vehicles, although this observation is not concrete. More research would need to be done on this subject that is not within the scope of this paper.

### Summary
Overall, by analyzing ride hail wait times we saw many of the same conclusions drawn by the ridership percentages verified. Ride hail service capabilities is affected significantly by the mode choice structure chosen. Initial plan, and whether or not they include ride hail, does affect service capabilities but only minimally (potential for longer wait times). Finally, the utility parameters used to calculate the mode choice does affect ridership. Ridership will increase when person and location variables are included. There is also potential for longer wait times when those variables are not included. These conclusions are clear, however, much of the results can be further explained by things beyond our control within this study. Section \@ref(deep-look) takes a deeper look at why some patterns in the ridership and wait times results exist.

## A Deeper Look at Ridership and Wait Times {#deep-look}
The results from Table \@ref(tab:tbexperiments) and the results from Figure \@ref(fig:waits) can be explained further by understanding the original setup of the experiments. For example, the clearest distinction in ridership and wait times exist between BEAM mode choice structures. The *None*, *RideHail*, and *All* structure types each produce results at a different magnitude. These vastly different results are directly related to how each model structure was constructed. To best explain the structure of each type however, Section \@ref(type1) fist explains the *None* structure, Section \@ref(type3) then explains the *All* structure, and finally Section \@ref(type2) explains the *RideHail* structure. The *All* structure is more complex than the *RideHail* structure, but the *RideHail* structure was constructed in part due to the problems within the *All* structure. For this reason, we explain the *All* structure before the *RideHail* structure. 

### BEAM (None) Effect on Results {#type1}
The *None* BEAM mode choice model type uses no mode choice code from the BEAM software. As explained in Table \@ref(tab:tbexperiments), this scenario descriptor points toward turning off BEAM's modal innovation. In a way, the results from these scenarios are almost identical to the direct output of ActivitySim itself. Since BEAM locked every trip's mode selection, we used these experiments to help understand the difference between how BEAM estimates ride hail against how ActivitySim estimates ride hail. Although it may seem illogical to use BEAM on top of ActivitySim for the *None* type scenarios, it was still necessary to run them with BEAM because BEAM estimates ride hail service capabilities with more detail than ActivitySim (See Section \@ref(lit-mas)).

### BEAM (All) Effect on Results {#type3}
The *All* BEAM mode choice model type uses a linked mode choice model with that of ActivitySim (See Section \@ref(meth-beam-link)) with all modal alternatives available. This adjusted model structure helps us better understand why we obtained much higher ridership than with the *None* type. Figure \@ref(fig:piechart) shows from which modes agents who switched to ride hail came from. Interestingly enough, the majority of agents who select ride hail switched from car type modes (car, hov2, and hov3). Adding together the car type modes, 74% of agents from the AsimBeamAll:Path scenario, 73% from the AsimBeamAll:PPL scenario, 74% of agents from the BeamAll:Path scenario, and 75% of agents from the BeamAll:PPL scenario  switch from a car type mode to a ride hail type mode by the last iteration. While about three-fourths of agents switch to ride hail from car type modes, 14-17% of agents switch from walk modes. Why is it that the *All* BEAM mode choice structure projects that the majority of ride hail users originate from car type modes? The *All* mode choice structure is consistent with that of ActivitySim, yet 75% of ride hail users go against ActivitySim's original mode choice selection to use a car type vehicle.

```{r piechart, echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "Original mode choices of agents who switch to a ride hail type mode.", fig.align='center'}
tar_read(piechart)
#planshifts <- tar_read(planshifts)
#ggsave("planshifts.png", width = 9, height = 6)
# below is for wRH-All-All scenario
```

We offer two factors as to why so many car users switch to ride hail modes. The first is that with the consistent mode choice structure, new utility parameter values are used to calculate the mode choice utility. The new utility parameters boost the ride hail utility, turning ride hail options into attractive alternatives. Figure \@ref(fig:sankey) provides sufficient evidence for this claim. Figure \@ref(fig:sankey) shows a sankey diagram of all modal decisions at the start of each iteration for the AsimBeamAll:PPL scenario. The mode "No Mode" describes those modes that were cleared and reset at the beginning of each iteration. (~10% of all trips had their modes reset each iteration). Notice how many of the car, hov2, hov2 passenger, hov3, and hov3 passenger modes shift into the "No Mode" category each iteration. Also notice in the subsequent iteration how many of those "No Mode" choices shift to ride hail modes! A shift from the "No Mode" choice to ride hail represents those agents choosing their mode based on the utility value.

\begin{landscape}

\begin{figure}[H]
\centering
\includegraphics{planshifts.png}
\caption{Sankey plot of cool stuff}
\label{fig:sankey}
\end{figure}

\end{landscape}

The second, more influencing factor for why agents switch from car to ride hail has to do with how BEAM is programmed. BEAM implements a complex car tracking algorithm which keeps track of household and agent level vehicle allocation. For example, if a household owns one vehicle, then the vehicle is *assigned* to the first agent to use it in the daily plan. This leaves the other household members to choose an alternate mode. ActivitySim monitors vehicle and agent level vehicle allocation differently by using auto sufficiency to estimate car mode choice at the household level. (Auto sufficient households are more likely to choose car than auto deficient households). BEAM's implementation of vehicular assignment prevents many agents from selecting a car mode, whereas ActivitySim may not prevent those same individuals from car usage. In addition to BEAM's vehicle assignment algorithm, BEAM's trip based mode choice structure also forces some car users to switch modes. Sometimes, agents will *lose* their vehicle within the day. This occurs when a pathway cannot be built. If an agent *loses* their car, they are likely to become unable to choose the car mode on future trips, as the car is usually gone and is no longer a valid modal alternative. Figure \@ref(fig:walkers) provides evidence for this statement. Figure \@ref(fig:walkers) displays a graph of those agents who start their day using a mode other than walk, but end up switching to the walk mode by the end of the day for the AsimBeamAll:PPL scenario. Notice how in each hour, as the percentage of walk users increases the  percentage of car type modes remains constant or decreases. This helps show that hundreds of car type mode users are switching to walk modes instead of remaining with their vehicle. Choosing walk on a car tour usually means the car path couldn't be built and the car is unable to be used on future trips within the day. Looking at Figure \@ref(fig:sankey) again, we notice that many walk users shift to ride hail each iteration. Therefore, many car type users who shift to walk during one trip, may shift to ride hail type vehicles on the next trip. This is true between corresponding iterations as well. 

```{r walkers, echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "Agents who switch to walk by time of day.", fig.align='center'}
#graph of those agents who start their day not walking, and switch to walking before the end of the day
#for wRH-All-All Scenario
tar_read(walk_switchers)
```

Overall, BEAM's internal code structure creates reasons to be unable to choose car modes, forcing agents to switch modal alternatives or get assigned the walk mode. The increase in ridership in the *All* model from the *None* model can be partially explained for these reasons.
 
### BEAM (RideHail) Effect on Results {#type2}
Finally, the way the *RideHail* BEAM mode choice model was constructed explains why their ridership and wait times were high. The *RideHail* model uses a linked mode choice model with that of ActivitySim, but only walk and transit users have the option to switch to ride hail type modes; car-type modes remained locked across each iteration. Whenever a ride hail path could be built, all walk modes were automatically given the option to choose ride hail or ride hail pooled and all transit modes were automatically given the option to choose ride hail transit. Although it made logical sense to lock all car-type modes (for reasons described in Section \@ref(type3)), by giving ride hail options only to walk and transit users, ridership increased even more than in the *All* scenario. The increase in ridership occurred because 1.) BEAM's adjusted code forced ride hail to be an option in almost all cases, and 2.) in most cases ride hail was calculated to be more attractive than walk or transit modes.

Table \@ref(tab:timeutil) provides evidence in ride hail being an attractive mode choice alternative. Table \@ref(tab:timeutil) displays the ride hail time utilization for each of the scenarios performed (except for  NoRideHail). The same ride hail fleet was used in each of the 9 scenarios. The ride hail fleet was composed of 952 ride hail driver shifts. Ride hail time utilization was calculated as the sum of all the driver shift times divided by the sum of all the passenger occupied ride travel time. Obviously, the AsimRideHail scenario had the lowest ride hail time utilization, at only 4.046%. This was because ActivitySim estimated a low ride hail ridership percentage and more drivers existed than riders. Interestingly, the *All* type scenarios ranged from 53.123% to 62.459% ride hail time utilization. This explains the higher wait times shown in Figure \@ref(fig:waits)! Higher ride hail utilization also explains why the Path models had higher maximum wait times. Finally, by analyzing the ride hail time utilization for the *RideHail* scenarios, we fully understand how attractive ride hail was. With 70.270% to 73.777% of ride hail time utilization present for the *RideHail* scenarios, we see three fourths of each driver's shift was used to transport passengers. This explains the attractiveness of the choice, the extreme increase in ridership, and  also the increased wait time for the *RideHail* type scenarios. In addition, it shows that if more driver shifts were available for the *RideHail* scenarios, wait times would have been lower and ridership would have been higher.  

```{r timeutil, echo=FALSE}
time_utilization <- tar_read(ridehail_utilization) %>% select(2,6)
kable(time_utilization, caption = 'Percent Ride Hail Time Utilization by Mode Choice Combination Scenario', booktabs = TRUE)
  #kable_styling(latex_options="scale_down")
```

```{r personutil, echo=FALSE}
#person_utilization <- tar_read(rh_utilization) %>% select(1,6)
#kable(person_utilization, caption = 'Ride hail (person / hour) utilization by mode choice combination scenario.', booktabs = TRUE)  %>%
#  kable_styling(latex_options="scale_down")
```

### Summary
As seen by the explanation of the structure of the *None*, *All*, and *RideHail* type scenarios, how BEAM's different mode choice structures are programmed affected total ride hail ridership and wait times significantly. By examining these structural differences, we better understand the effect mode choice has on forecasting novel modes. For example, the use of path, person, and location type variables cause ride hail to be a more attractive choice than if it were only calculated with the default BEAM parameters (See Section \@ref(novel-beam)). In addition, BEAM's advanced car tracking algorithm and trip based mode structure causes many car type trips to shift to ride hail after first defaulting to the walk mode. Finally, we conclude that the *RideHail* type scenario is best at forecasting novel mode usage, as shown by the high ride hail time utilization percentages. The *RideHail* type scenario also avoids the BEAM pitfalls of loosing car modes and provides ride hailing vehicles to those who need it most. Although this is true, its limitation exists in that *all* walk trips are given the option to shift to ride hail. In reality, only those walk trips who get their mode choice cleared after each iteration should have the opportunity to switch modes. This was not accounted for in the research methodology because of the complexity of the BEAM code adjustment. Section \@ref(discussion) further describes the deeper meaning behind the ride hail, wait time, and mode choice structure results discovered in this section.
